MultiATSM package - General Guidelines

### 2025-10-23 inst/doc/MultiATSM.Rmd
version 1.5.0

Due to peculiar features of JPS-based specifications, an efficient
estimation of the parameters governing the risk-neutral ($Q$-measure)
and the physical ($P$-measure) probability measures can be carried out
rather independently. The only exception is the variance-covariance
matrix (sigma) term which is a common element to both the $P$ and the
$Q$ density functions. In all cases, the risk factor dynamics under the
$P$-measure follow a VAR(1) model of some sort. The table below
summarizes the general features of each one of the models available at
this package.

```{r ModFea, message=FALSE, echo=FALSE}
ModelLabels <- c("JPS original", "JPS global", "JPS multi", "GVAR single", "GVAR multi", 
                 "JLL original", "JLL No DomUnit", "JLL joint Sigma")

@@

# Adjust column names
ColNames <- c("","","","","JLL", "GVAR", "", "", "", "", "", "", "","", "", "","")
colnames(Tab) <- ColNames

@@

kableExtra::kbl(Tab, align = "c", caption = "Model Features") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 10) %>%
  kableExtra::add_header_above(c(" "=2, "Unrestricted" = 1, "Restricted" = 1, "Unrestricted" = 1, "Restricted" = 2, " " = 11)) %>%
 kableExtra::add_header_above(c(" "=2, "Individual" = 2, "Joint" = 3, " "=2, "Individual" = 1, "Joint" = 1, " "=2, "P only" = 1, "P and Q" = 1, " " = 3))  %>%
  kableExtra::add_header_above(c( " "=2, "P-dynamics"= 5, " "=2, "Q-dynamics"= 2, " "=2, "Sigma matrix estimation" = 2, " "=2, "Dominant Country"=1), bold = T) %>%
kableExtra::pack_rows("Unrestricted VAR", 1, 3 , label_row_css = "background-color: #666; color: #fff;")  %>%
kableExtra::pack_rows("Restricted VAR (GVAR)", 4, 5, label_row_css = "background-color: #666; color: #fff;") %>%
kableExtra::pack_rows("Restricted VAR (JLL)", 6, 8, label_row_css = "background-color: #666; color: #fff;")
```

@@

In what follows, the focus will be devoted to describe the various
pieces that are necessary to implement ATSMs. Specifically, Section
\@ref(S:SectionData) describes the data-set available at this package
and the set of functions that are useful to retrieve data from Excel
files. Section \@ref(S:SectionInputs) exposes the necessary inputs that
have to be specified by the user. In Section \@ref(S:SectionEstimation)
the estimation procedure is detailed. Section
\@ref(S:SectionReplication) shows how to use the *MultiATSM* package to
estimate ATSMs from scratch.

# Data {#S:SectionData}

@@

```{r}
data('CM_Yields')
```

Bond yield data are also used to construct the country-specific **spanned** factors,
according to the terminology of JPS (2014). These factors typically correspond to 
the standard yield curve components, namely level, slope, and curvature. An intuitive 
explanation of their statistical properties is provided in Section \@ref(S:SpaFac).

@@

```{r}
data("CM_GlobalMacroFactors")
data('CM_DomMacroFactors')
```

@@

```{r}
data('CM_Trade')
```

@@

This package also offers an automatized procedure to extract data from
Excel files and to, subsequentially, prepare the risk factor databases
that are directly used in the model estimation. The use of the package
functions requires that the databases *(i)* are constructed in separate
Excel files for the unspanned factors, the term structure data, and the
measures of interdependence (for the GVAR-based models); *(ii)* contain,
in each Excel file, one separate tab per country (in addition to the
global variables, in the case of the unspanned factors' database); and
*(iii)* have identical variable labels across all the tabs within each
file. For illustration, see the Excel file available at the package. One
example of list of inputs to be provided is, for instance,

```{r}
Initial_Date <- "2006-09-01" # Format "yyyy-mm-dd"
Final_Date <- "2019-01-01" # Format "yyyy-mm-dd"
DataFrequency <- "Monthly"
GlobalVar <- c("GBC", "VIX") # Global Variables
DomVar <- c("Eco_Act", "Inflation", "Com_Prices", "Exc_Rates") #  Domestic variables
N <-  3 # Number of spanned factors per country
Economies <- c("China", "Mexico", "Uruguay", "Brazil", "Russia")
ModelType <- "JPS original"
```

Using these inputs, the variable *RiskFac_TS* can be constructed,
containing the full time series of risk factors required for the model.

```{r}
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
RiskFac_TS <- DataForEstimation(Initial_Date, Final_Date, Economies, N, FactorLabels, ModelType,
                                DataFrequency)
```

# Required user inputs {#S:SectionInputs}

@@

-   *Model type*: a character vector indicating the model type to be
    estimated, as presented in Table \@ref(tab:ModFea).

@@

-   A binary variable that equals $1$ if the user wishes to impose the
    constraint that the largest eigenvalue under the $Q$-measure is
    strictly less than $1$ otherwise, it should be set to $0$. While
    enforcing this stationary constraint may prolong model estimation,
    it can improve the convergence of the model.

**6. Output label**

@@

# 5) Risk-neutral stationary constraint
StationarityUnderQ <- 0 # 1 = set stationary condition under the Q; 0 = no stationary condition

# 6) Output label
OutputLabel <- "Model_demo"
```

@@

Additional inputs are required for estimating GVAR-ATSM or JLL-ATSM
models. These inputs should be organized into separate lists, named
*GVARlist* and *JLLlist* for clarity. This section outlines the general
structure of both lists, while Sections \@ref(S:SectionGVAR) and
\@ref(S:SectionJLL) provide detailed explanations and available options.

@@

```{r}
W_type <- 'Sample Mean' # Method to compute the transition matrix
t_First_Wgvar <- "2000" # First year of the sample
t_Last_Wgvar <-  "2015" # Last year of the sample
DataConnectedness <- TradeFlows # Measure of connectedness across countries
```

@@

```{r}
GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2000",
                   t_Last_Wgvar = "2015", DataConnectedness = TradeFlows) 
```

@@

-   *Mean or median of OLS estimates* *(flag_mean)*: the user must
    decide whether to compute the mean or median of the OLS estimates
    after each bootstrap iteration. To compute the mean (median), the
    user must set *TRUE* (*FALSE*);

@@

```{r}
BRWlist <- within(list(flag_mean = TRUE, gamma = 0.2, N_iter = 500, B = 50, 
                       checkBRW = TRUE, B_check = 1000), N_burn <- round(N_iter * 0.15))
```

@@

ATSMInputs <- InputsForOpt(t0, tF, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,  
                           FactorLabels, Economies, DataFreq, CheckInputs = FALSE)
```

@@

```{r}
WishGraphRiskFac <- 0 #   YES: 1; No = 0.
WishGraphYields <- 1 #    YES: 1; No = 0.
WishOrthoJLLgraphs <- 0 # YES: 1; No = 0.
```

The function `InputsForOutputs()` can provide some guidance for
customizing the features of the wished outputs. Conditional to these
settings, individual folders are created at the user's *temporary
directory* (which can be accessed using `tempdir()` ) to store the
different types of the desired graphical outputs.

@@

In its current form, the *MultiATSM* package allows for the calculation
of two risk compensation measures: term premia and forward premia. In
particular, a $n$-maturity bond yield can be decomposed as the sum of
its expected short-rate ($Exp_t^{(n)}$) and term premia 
($TP_t^{(n)}$) components.
Technically: $$
y_t^{(n)} = Exp_t^{(n)} + TP_t^{(n)}
$$ In the package's standard form, the expected short rate term is
computed from time $t$ to $t+n$, which represents the bond's maturity:
$Exp_t^{(n)} = \sum_{h=0}^{n} E_t[y_{t+h}^{(1)}]$. Alternatively, one
could perfom such a decomposition at the forward rate level
($f_t^{(n)}$), specifically
$f_t^{(n)} = \sum_{h=m}^{n} E_t[y_{t+h}^{(1)}] + FP_t^{(n)}$ where
$FP_t^{(n)}$ corresponds to the forward premia. In this case, the user
must specify a binary variable set to $1$ if the computation of forward
premia is desired, or $0$ otherwise. If set to $1$, the user must also
provide a two-element numerical vector containing the maturities
corresponding to the starting and ending dates of the bond maturity.
Example:

```{r}
    WishFPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0 
    FPmatLim <- c(60, 120)
```

@@

```{r}
Bootlist <- list(methodBS = 'block', BlockLength = 4, ndraws =  50, pctg   =  95)
```

@@

# Model Estimation {#S:SectionEstimation}

@@

## Spanned Factors {#S:SpaFac}

@@

w_pca <- data.frame(t(w[1:N,]))
colnames(w_pca) <- LabSpaFac
w_pca$mat <- c(0.25, 0.5, 1, 3, 5, 10) # vector of maturitie

@@

g <-  ggplot2::ggplot(data = w_pca, ggplot2::aes(x=  mat)) +  
    ggplot2::geom_line(ggplot2::aes(y = Level, color = "Level"), linewidth = 0.7) + 
    ggplot2::geom_line(ggplot2::aes(y = Slope, color = "Slope"), linewidth = 0.7) +
    ggplot2::geom_line(ggplot2::aes(y = Curvature, color = "Curvature"),  linewidth = 0.7) +
      ggplot2::labs(color = "Legend") + ggplot2::scale_color_manual(values = colors) + ggplot2::theme_classic() +
    ggplot2::theme(axis.title.y= ggplot2::element_blank(), legend.position="top", legend.title=ggplot2::element_blank(), legend.text= ggplot2::element_text(size=8) ) + 
   ggplot2::xlab("Maturity (Years)") + ggplot2::geom_hline(yintercept=0)

@@

```{r}
data('CM_Yields')
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
N <- 3
SpaFact <- Spanned_Factors(Yields, Economies, N)
```

@@

As presented in Table \@ref(tab:ModFea), the dynamics of the risk
factors under the $P$-measure evolve according to a VAR(1) model, which
may be fully unrestricted (as in the case of the JPS-related models) or
partially restricted (as in the GVAR and JLL frameworks). While the
focus here is their integration into an ATSM setup, each specification
can also be estimated independently to study dynamic systems. The usage
of these models is illustrated below.

@@

```{r}
data("CM_Factors")
PdynPara <- VAR(RiskFactors, VARtype= "unconstrained")
```

@@

```{r}
FactorsChina <- RiskFactors[1:7, ]
PdynPara <- VAR(FactorsChina, VARtype= "unconstrained")
```

@@

### GVAR {#S:SectionGVAR}

@@

```{r}
data('CM_Factors_GVAR')
GVARinputs <- list(Economies = Economies, GVARFactors = FactorsGVAR, VARXtype ="constrained: Inflation")
```

@@

```{r}
data("CM_Trade")
t_First <- "2006"
t_Last <-  "2019"
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
type <- "Sample Mean"
W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, DataPath = NULL, TradeFlows)
print(W_gvar)
```

@@

```{r}
data("CM_Factors_GVAR")
GVARinputs <- list(Economies = Economies, GVARFactors = FactorsGVAR, VARXtype = "unconstrained", 
                   Wgvar = W_gvar)
N <- 3

@@

A separate routine is provided for computing the foreign-specific
factors (also commonly referred to as the star variables) used in the
estimation of the VARX models.

```{r}
data('CM_Factors')
StaFac <- StarFactors(RiskFactors, Economies, W_gvar)
```

### JLL {#S:SectionJLL}

@@

3.  *Wish the estimation of the sigmas matrices*: this is a binary
    variable which assumes value equal to $1$ if the user wishes the
    estimation of all JLL sigma matrices (i.e. variance-covariance and
    the Cholesky factorization matrices) and, $0$ otherwise. Since this
    estimation is done numerically, its estimation can take several
    minutes;

@@

5.  *JLL type*: a string-vector containing the label of the JLL model to
    be estimated according to the features described in Table
    \@ref(tab:ModFea).

@@

```{r}
ModelType <- "JLL original"   
JLLinputs <- list(Economies = Economies, DomUnit = "China", WishSigmas = 1,  SigmaNonOrtho = NULL,
                  JLLModelType = ModelType)
```

@@

```{r, eval=FALSE}
data("CM_Factors")
N <- 3
JLLpara <- JLL(RiskFactors, N, JLLinputs, CheckInputs = TRUE)
```

@@

ATSM estimation involves optimizing the log-likelihood function by
optimally selecting the model parameters. The optimization structure in
this package is based on routines from the term structure package by
@LeSingleton2018. The ATSM otimization is carried out using the
`Optimization()` function.

Broadly, the unspanned risk framework from JPS (2014) (and, therefore,
all its multi-country extensions) requires the estimation parameter set
consisting of 6 elements, namely: the risk-neutral long-run mean of the
short rate (*r0*), the risk-neutral feedback matrix (*K1XQ*), the
variance-covariance matrix (*SSZ*) from the VAR processes, the standard
deviation of the errors from the portfolios of yields observed with
error (*se*), in addition to the intercept (*K0Z*) and the feedback
(*K1Z*) matrices of the physical dynamics.

The parameters *K0Z* and *K1Z* are estimated as part of the model's
$P-$dynamics and are known in closed form. Likewise, *r0* and *se* have
analytical solutions and can be factored out of the log-likelihood
function. However, the remaining parameters (*K1XQ* and *SSZ*) require
numerical estimation. @LeSingleton2018 provide standardized routines
(integrated into this package) to set effective initial values,
facilitating the optimization process.

The *MultiATSM* package offers two optimization algorithms: *fminunc*
and *fminsearch*. By default, both algorithms are run sequentially twice
to ensure the model reaches the global optimum. Specifically in the
default bootstrap setup, only *fminunc* is used to save time.

@@

This section presents some examples on how to use the *MultiATSM*
package to fully implement ATSMs. The console displays messages to help
the user track the progress of the code execution at each step. Further
on, the functions of this package are used to replicate some of the
results presented in the original papers of JPS (2014), CM (2023) and CM
(2024). See the [Paper Replication Vignette](Paper-Replications.html) of
this package for a detailed reproduction.

@@

```{r, eval=FALSE}
########################################################################################################
#################################### USER INPUTS #######################################################
########################################################################################################
# A) Load database data
LoadData("CM_2024")

# B) GENERAL model inputs
ModelType <- "JPS multi" # available options: "JPS original", "JPS global", "GVAR single", "JPS multi",
                          #"GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".

Economies <- c("China", "Brazil", "Mexico") # Names of the economies from the economic system
GlobalVar <- c("Gl_Eco_Act")# Global Variables
DomVar <- c("Eco_Act", "Inflation") # Country-specific variables
N <- 2  # Number of spanned factors per country

t0_sample <- "01-05-2005" # Format: "dd-mm-yyyy"
tF_sample <- "01-12-2019" # Format: "dd-mm-yyyy"

OutputLabel <- "Test" # label of the model for saving the file
DataFreq <-"Monthly" # Frequency of the data

StatQ <- 0 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1,NO:0

# B.1) SPECIFIC model inputs
#################################### GVAR-based models ##################################################
GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2005",
                  t_Last_Wgvar = "2019", DataConnectedness <- TradeFlows ) 
# VARXtype: Available options "unconstrained" or "constrained" (VARX)
# W_type: Method to compute the transition matrix. Options:"Time-varying" or "Sample Mean"
# t_First_Wgvar: First year of the sample (transition matrix)
# t_Last_Wgvar:  Last year of the sample (transition matrix)
# DataConnectedness: measure of connectedness across countries
#################################### JLL-based models ###################################################
JLLlist <- list(DomUnit =  "China")
# DomUnit: name of the economy of the economic system, or "None" for the model "JLL No DomUnit"
###################################### BRW inputs  ######################################################
WishBC <- 0 # Wish to estimate the model with the bias correction method of BRW (2012): #YES: 1, NO:0
BRWlist <- within(list(flag_mean = TRUE, gamma = 0.05, N_iter = 250, B = 50, checkBRW = TRUE,
                       B_check = 1000),  N_burn <- round(N_iter * 0.15))
# flag_mean: TRUE = compute the mean; FALSE = compute the median
# gamma: Adjustment parameter
# N_iter:  Number of iteration to be conserved
# N_burn:  Number of iteration to be discarded
# B: Number of bootstrap samples
# checkBRW: wishes to perform closeness check: TRUE or FALSE
# B_check: If checkBRW is chosen as TRUE, then choose number of bootstrap samples used in the check

# C) Decide on Settings for numerical outputs
WishFPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0
FPmatLim <- c(60,120) #  If the forward premia is desired, then choose the Min and max maturities of the
                      # forward premia. Otherwise set NA
Horiz <- 30
DesiredGraphs <- c("Fit", "IRF", "TermPremia") # "Fit", "IRF", "FEVD", "GIRF", "GFEVD", "TermPremia"
WishGraphRiskFac <- 0
WishGraphYields <- 1
WishOrthoJLLgraphs <- 0

# D) Bootstrap settings
WishBootstrap <- 1 #  YES: 1; No = 0.
BootList <- list(methodBS = 'bs', BlockLength = 4, ndraws = 5, pctg =  95)
# methodBS: bootstrap method. Available options: (i) 'bs' ; (ii) 'wild'; (iii) 'block'
# BlockLength: Block length, necessary input for the block bootstrap method
# ndraws: number of bootstrap draws
# pctg: confidence level

>>>>

Due to peculiar features of JPS-based specifications (see @JoslinSingletonZhu2011), an efficient
estimation of the parameters governing the risk-neutral ($Q$-measure)
and the physical ($P$-measure) probability measures can be carried out
rather independently. The only exception is the variance-covariance
matrix (sigma) term which is a common element to both the $P$ and the
$Q$ density functions. In all cases, the risk factor dynamics under the
$P$-measure follow a VAR(1) model of some sort. The table below
summarizes the general features of each one of the models available at
this package.

<a id="tab-ModFea"></a>

```{r ModFea, message=FALSE, echo=FALSE}
ModelLabels <- c(
  "JPS original", "JPS global", "JPS multi", "GVAR single", "GVAR multi",
  "JLL original", "JLL No DomUnit", "JLL joint Sigma"
)

@@

# Adjust column names
ColNames <- c("", "", "", "", "JLL", "GVAR", "", "", "", "", "", "", "", "", "", "", "")
colnames(Tab) <- ColNames

@@

kableExtra::kbl(
  Tab,
  align = "c",
  caption = "Model Features"
) %>%
  kableExtra::kable_classic("striped", full_width = F) %>%
  kableExtra::row_spec(0, font_size = 10) %>%
  kableExtra::add_header_above(c(" " = 2, "Unrestricted" = 1, "Restricted" = 1, "Unrestricted" = 1, "Restricted" = 2, " " = 11)) %>%
  kableExtra::add_header_above(c(" " = 2, "Individual" = 2, "Joint" = 3, " " = 2, "Individual" = 1, "Joint" = 1, " " = 2, "P only" = 1, "P and Q" = 1, " " = 3)) %>%
  kableExtra::add_header_above(c(" " = 2, "P-dynamics" = 5, " " = 2, "Q-dynamics" = 2, " " = 2, "Sigma matrix estimation" = 2, " " = 2, "Dominant Country" = 1), bold = T) %>%
  kableExtra::pack_rows("Unrestricted VAR", 1, 3, label_row_css = "background-color: #666; color: #fff;") %>%
  kableExtra::pack_rows("Restricted VAR (GVAR)", 4, 5, label_row_css = "background-color: #666; color: #fff;") %>%
  kableExtra::pack_rows("Restricted VAR (JLL)", 6, 8, label_row_css = "background-color: #666; color: #fff;")
```

@@

In what follows, the focus will be devoted to describe the various
pieces that are necessary to implement ATSMs. Specifically, [Section
Data](#SectionData) describes the data-set available at this package
and the set of functions that are useful to retrieve data from Excel
files. [Section Required user inputs](#SectionInputs) exposes the necessary inputs that
have to be specified by the user. In Section [Model Estimation](#SectionEstimation)
the estimation procedure is detailed. Section
[Model Estimation](#SectionEstimation) shows how to use the *MultiATSM* package to
estimate ATSMs from scratch.

# Data {#SectionData}

@@

```{r}
data("Yields")
```

Bond yield data are also used to construct the country-specific **spanned** factors,
according to the terminology of JPS (2014). These factors typically correspond to 
the standard yield curve components, namely level, slope, and curvature. An intuitive 
explanation of their statistical properties is provided in Section [Spanned Factors](#SpaFac).

@@

```{r}
data("GlobalMacro")
data("DomMacro")
```

@@

```{r}
data("TradeFlows")
```

@@

This package also offers an automatized procedure to extract data from
Excel files and to, subsequentially, prepare the risk factor databases
that are directly used in the model estimation. The use of the package
functions requires that the databases *(i)* are constructed in separate
Excel files for the unspanned factors, the term structure data, and the
measures of interdependence (for the GVAR-based models); *(ii)* contain,
in each Excel file, one separate tab per country (in addition to the
global variables, in the case of the unspanned factors' database); and
*(iii)* have identical variable labels across all the tabs within each
file. For illustration, see the Excel file available at the package. 
The workflow is as follows:

  1. Retrieve the full dataset from excel files using `Load_Excel_Data()`:
```{r}
MacroData <- Load_Excel_Data(system.file("extdata", "MacroData.xlsx", package = "MultiATSM"))
YieldData <- Load_Excel_Data(system.file("extdata", "YieldsData.xlsx", package = "MultiATSM"))
```

  2. Specify a subset of the selected inputs. For instance,
```{r}
Initial_Date <- "2006-09-01" # Format "yyyy-mm-dd"
Final_Date <- "2019-01-01" # Format "yyyy-mm-dd"
DataFrequency <- "Monthly"
GlobalVar <- c("GBC", "VIX") # Global Variables
DomVar <- c("Eco_Act", "Inflation", "Com_Prices", "Exc_Rates") #  Domestic variables
N <- 3 # Number of spanned factors per country
Economies <- c("China", "Mexico", "Uruguay", "Brazil", "Russia")
ModelType <- "JPS original"
```

  3. Conditional on the previous step, the variable *RiskFac_TS* can be constructed. This variable
contains the full time series of risk factors required for the model.

```{r}
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
RiskFac_TS <- DataForEstimation(
  Initial_Date, Final_Date, Economies, N, FactorLabels, ModelType,
  DataFrequency, MacroData, YieldData
)
```

# Required user inputs {#SectionInputs}

@@

-   *Model type*: a character vector indicating the model type to be
    estimated, as presented in [Table&nbsp;0.1](#tab-ModFea).

@@

-   A logical variable. Choose *TRUE* to impose the
    constraint that the largest eigenvalue under the $Q$-measure is
    strictly less than $1$ otherwise, it should be set to *FALSE*. While
    enforcing this stationary constraint may prolong model estimation,
    it can improve the convergence of the model.

**6. Selected folder to save outputs**

-  Path where the selected graphical outputs will be saved. If set to `NULL`, the outputs are stored in the user's *temporary directory* (accessible via `tempdir()`).

**7. Output label**

@@

# 5) Risk-neutral stationary constraint
StationarityUnderQ <- FALSE

# 6) Path to save outputs
Folder2Save <- NULL

# 7) Output label
OutputLabel <- "Model_demo"
```

@@

Additional inputs are required for estimating GVAR-ATSM or JLL-ATSM
models. These inputs should be organized into separate lists, named
*GVARlist* and *JLLlist* for clarity. This section outlines the general
structure of both lists, while Sections [GVAR](#SectionGVAR) and
[JLL](#SectionJLL) provide detailed explanations and available options.

@@

```{r}
W_type <- "Sample Mean" # Method to compute the transition matrix
t_First_Wgvar <- "2000" # First year of the sample
t_Last_Wgvar <- "2015" # Last year of the sample
DataConnectedness <- TradeFlows # Measure of connectedness across countries
```

@@

```{r}
GVARlist <- list(
  VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2000",
  t_Last_Wgvar = "2015", DataConnectedness = TradeFlows
)
```

@@

-   *Mean or median of P-dynamics estimates* *(Cent_Measure)*: the user must
    decide whether to compute the mean or median of the physical dynamics estimates
    after each bootstrap iteration. To compute the mean (median), the
    user must set *Mean* (*Median*);

@@

-   *Eigenvalue restriction (Eigen_rest)*: set a restriction on the largest eigenvalue 
     under the $P$-measure. Default is $1$.

```{r}
BRWlist <- within(list(
  Cent_Measure = "Mean", gamma = 0.2, N_iter = 500, B = 50,
  checkBRW = TRUE, B_check = 1000, Eigen_rest = 1
), N_burn <- round(N_iter * 0.15))
```

@@

ATSMInputs <- InputsForOpt(t0, tF, ModelType, Yields, GlobalMacro, DomMacro,
  FactorLabels, Economies, DataFreq,
  CheckInputs = FALSE
)
```

@@

```{r}
WishGraphRiskFac <- FALSE
WishGraphYields <- TRUE
WishOrthoJLLgraphs <- FALSE
```

The function `InputsForOutputs()` can provide some guidance for
customizing the features of the wished outputs. Conditional to these
settings, individual folders are created at the selected `Folder2Save` to store the
different types of the desired graphical outputs. The selected outputs are also printed to the user' console.

@@

In its current form, the *MultiATSM* package allows for the calculation
of two risk compensation measures: term premia and forward premia. In
particular, a $n$-maturity bond yield can be decomposed as the sum of
its expected short-rate ($Exp_t^{(n)}$) and term premia 
($TP_t^{(n)}$) components.
Technically: $$
y_t^{(n)} = Exp_t^{(n)} + TP_t^{(n)}
$$ In the package's standard form, the expected short rate term is
computed from time $t$ to $t+n$, which represents the bond's maturity:
$Exp_t^{(n)} = \sum_{h=0}^{n} E_t[y_{t+h}^{(1)}]$. Alternatively, one
could perfom such a decomposition at the forward rate level
($f_t^{(n)}$), specifically
$f_t^{(n)} = \sum_{h=m}^{n} E_t[y_{t+h}^{(1)}] + FP_t^{(n)}$ where
$FP_t^{(n)}$ corresponds to the forward premia. In this case, the user
must specify a logical variable set to *TRUE* if the computation of forward
premia is desired, or *FALSE* otherwise. If set to *TRUE*, the user must also
provide a two-element numerical vector containing the maturities
corresponding to the starting and ending dates of the bond maturity.
Example:

```{r}
WishFPremia <- TRUE
FPmatLim <- c(60, 120)
```

@@

```{r}
Bootlist <- list(methodBS = "block", BlockLength = 4, ndraws = 50, pctg = 95)
```

@@

# Model Estimation {#SectionEstimation}

@@

## Spanned Factors {#SpaFac}

@@

w_pca <- data.frame(t(w[1:N, ]))
colnames(w_pca) <- LabSpaFac
w_pca$mat <- c(0.25, 0.5, 1, 3, 5, 10) # vector of maturitie

@@

g <- ggplot2::ggplot(data = w_pca, ggplot2::aes(x = mat)) +
  ggplot2::geom_line(ggplot2::aes(y = Level, color = "Level"), linewidth = 0.7) +
  ggplot2::geom_line(ggplot2::aes(y = Slope, color = "Slope"), linewidth = 0.7) +
  ggplot2::geom_line(ggplot2::aes(y = Curvature, color = "Curvature"), linewidth = 0.7) +
  ggplot2::labs(color = "Legend") +
  ggplot2::scale_color_manual(values = colors) +
  ggplot2::theme_classic() +
  ggplot2::theme(axis.title.y = ggplot2::element_blank(), legend.position = "top", legend.title = ggplot2::element_blank(), legend.text = ggplot2::element_text(size = 8)) +
  ggplot2::xlab("Maturity (Years)") +
  ggplot2::geom_hline(yintercept = 0)

@@

```{r}
data("Yields")
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
N <- 3
SpaFact <- Spanned_Factors(Yields, Economies, N)
```

@@

As presented in [Table&nbsp;0.1](#tab-ModFea), the dynamics of the risk
factors under the $P$-measure evolve according to a VAR(1) model, which
may be fully unrestricted (as in the case of the JPS-related models) or
partially restricted (as in the GVAR and JLL frameworks). While the
focus here is their integration into an ATSM setup, each specification
can also be estimated independently to study dynamic systems. The usage
of these models is illustrated below.

@@

```{r}
data(RiskFacFull)
PdynPara <- VAR(RiskFacFull, VARtype = "unconstrained")
```

@@

```{r}
FactorsChina <- RiskFacFull[1:7, ]
PdynPara <- VAR(FactorsChina, VARtype = "unconstrained")
```

@@

### GVAR {#SectionGVAR}

@@

```{r}
data(GVARFactors)
GVARinputs <- list(Economies = Economies, GVARFactors = GVARFactors, VARXtype = "constrained: Inflation")
```

@@

```{r}
data("TradeFlows")
t_First <- "2006"
t_Last <- "2019"
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
type <- "Sample Mean"
W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, TradeFlows)
print(W_gvar)
```

@@

```{r}
data("GVARFactors")
GVARinputs <- list(
  Economies = Economies, GVARFactors = GVARFactors, VARXtype = "unconstrained",
  Wgvar = W_gvar
)
N <- 3

@@

### JLL {#SectionJLL}

@@

3.  *Wish the estimation of the sigmas matrices*: this is a logical
    variable which assumes value equal to *TRUE* if the user wishes the
    estimation of all JLL sigma matrices (i.e. variance-covariance and
    the Cholesky factorization matrices) and, *FALSE* otherwise. Since this
    estimation is done numerically, its estimation can take several
    minutes;

@@

5.  *JLL type*: a string-vector containing the label of the JLL model to
    be estimated according to the features described in [Table&nbsp;0.1](#tab-ModFea).

@@

```{r}
ModelType <- "JLL original"
JLLinputs <- list(
  Economies = Economies, DomUnit = "China", WishSigmas = TRUE, SigmaNonOrtho = NULL,
  JLLModelType = ModelType
)
```

@@

```{r, eval=FALSE}
data("RiskFacFull")
N <- 3
JLLpara <- JLL(RiskFacFull, N, JLLinputs, CheckInputs = TRUE)
```

@@

ATSM estimation involves optimizing the log-likelihood function by
optimally selecting the model parameters via the `Optimization()` function.
Broadly, the unspanned risk framework from JPS (2014) (and, therefore,
all its multi-country extensions) follows a model parameterization similar to that proposed by @JoslinSingletonZhu2011. Specifically, it requires the estimation parameter set
consisting of 6 elements, namely the:

1.    Risk-neutral long-run mean of the
short rate (*r0*);

2.    Risk-neutral feedback matrix (*K1XQ*);

3.    Variance-covariance matrix (*SSZ*) from the VAR processes;

4.    The standard deviation of the errors from the portfolios of yields observed with
error (*se*);

5.    Intercept of the physical dynamics (*K0Z*);

6.    Feedback matrices of the physical dynamics (*K1Z*).

The parameters *K0Z* and *K1Z* are estimated as part of the model's
$P-$dynamics and are known in closed form. Likewise, *r0* and *se* have
analytical solutions and can be factored out of the log-likelihood
function. However, the remaining parameters (*K1XQ* and *SSZ*) require
numerical estimation. This package provides standardized routines
to set effective initial values, facilitating the optimization process.

The *MultiATSM* package offers two optimization algorithms: *L-BFGS-B* and *Nelder-Mead*. By default, both algorithms are run sequentially twice
to ensure the model reaches the global optimum. Specifically in the
default bootstrap setup, only *Nelder-Mead* is used to save time.

@@

This section presents some examples on how to use the *MultiATSM*
package to fully implement ATSMs. The console displays messages to help
the user track the progress of the code execution at each step. Further
on, the functions of this package are used to replicate some of the
results presented in the original papers of JPS (2014), CM (2023) and CM
(2024). See the **paper replication vignette** (included with this package) for a detailed reproduction.

@@

```{r, eval=FALSE}
########################################################################################################
#################################### USER INPUTS #######################################################
########################################################################################################
library(MultiATSM)
# A) Load database data
LoadData("CM_2024")

# B) GENERAL model inputs
ModelType <- "JPS original" # available options: "JPS original", "JPS global", "GVAR single", "JPS multi",
# "GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".

Economies <- c("China", "Brazil") # Names of the economies from the economic system
GlobalVar <- c("Gl_Eco_Act") # Global Variables
DomVar <- c("Eco_Act") # Country-specific variables
N <- 2 # Number of spanned factors per country

t0_sample <- "01-12-2007" # Format: "dd-mm-yyyy"
tF_sample <- "01-12-2019" # Format: "dd-mm-yyyy"

OutputLabel <- "Test" # label of the model for saving the file
DataFreq <- "Monthly" # Frequency of the data

Folder2Save <- NULL # Folder to save the required graphical outputs. NULL -> temporary directory
StatQ <- FALSE # Wish to impose stationary condition for the eigenvalues of each country

# B.1) SPECIFIC model inputs
#################################### GVAR-based models ##################################################
GVARlist <- list(
  VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2005",
  t_Last_Wgvar = "2019", DataConnectedness = TradeFlows
)
# VARXtype: Available options "unconstrained" or "constrained" (VARX)
# W_type: Method to compute the transition matrix. Options:"Time-varying" or "Sample Mean"
# t_First_Wgvar: First year of the sample (transition matrix)
# t_Last_Wgvar:  Last year of the sample (transition matrix)
# DataConnectedness: measure of connectedness across countries
#################################### JLL-based models ###################################################
JLLlist <- list(DomUnit = "China")
# DomUnit: name of the economy of the economic system, or "None" for the model "JLL No DomUnit"
###################################### BRW inputs  ######################################################
WishBC <- FALSE # Wish to estimate the model with the bias correction method of BRW (2012)
BRWlist <- within(list(
  Cent_Measure = "Mean", gamma = 0.05, N_iter = 250, B = 50, checkBRW = TRUE,
  B_check = 1000, Eigen_rest = 1
), N_burn <- round(N_iter * 0.15))
# Cent_Measure: "Mean" = compute the mean; "Median" = compute the median
# gamma: Adjustment parameter
# N_iter:  Number of iteration to be conserved
# N_burn:  Number of iteration to be discarded
# B: Number of bootstrap samples
# checkBRW: wishes to perform closeness check: TRUE or FALSE
# B_check: If checkBRW is chosen as TRUE, then choose number of bootstrap samples used in the check

# C) Decide on Settings for numerical outputs
WishFPremia <- TRUE # Wish to estimate the forward premia
FPmatLim <- c(60, 120) #  If the forward premia is desired, then choose the Min and max maturities of the
# forward premia. Otherwise set NA
Horiz <- 30
DesiredGraphs <- c("Fit", "IRF", "TermPremia") # "Fit", "IRF", "FEVD", "GIRF", "GFEVD", "TermPremia"
WishGraphRiskFac <- FALSE
WishGraphYields <- TRUE
WishOrthoJLLgraphs <- FALSE

# D) Bootstrap settings
WishBootstrap <- TRUE
BootList <- list(methodBS = "bs", BlockLength = 4, ndraws = 4, pctg = 95)
# methodBS: bootstrap method. Available options: (i) 'bs' ; (ii) 'wild'; (iii) 'block'
# BlockLength: Block length, necessary input for the block bootstrap method
# ndraws: number of bootstrap draws
# pctg: confidence level
